Scaling through abstractions – high-performance vectorial wave simulations for seismic inversion with Devito

### 2020-04-21 paper.tex
final

\title{Scaling through abstraction – high-performance vectorial wave simulations for seismic inversion with Devito}

\author{
\IEEEauthorblockN{1\textsuperscript{st} Mathias Louboutin}
\IEEEauthorblockA{\textit{Georgia Institute of Technology}\\
Antlanta, GA \\
mlouboutin3@gatech.edu}
\and
\IEEEauthorblockN{2\textsuperscript{nd} Fabio Luporini}
\IEEEauthorblockA{\textit{Devito Codes}\\
London, UK \\
fabio@devitocodes.com}
\and
\IEEEauthorblockN{3\textsuperscript{th} Rhodri Nelson}
\IEEEauthorblockA{\textit{Imperial College London}\\
London, UK \\
rhodri.nelson@imperial.ac.uk}
\and
\IEEEauthorblockN{4\textsuperscript{rd} Philipp Witte}
\IEEEauthorblockA{\textit{Georgia Institute of Technology}\\
Atlanta, GA \\
pwitte3@gatech.edu}
\and
\IEEEauthorblockN{5\textsuperscript{th} George Bisbas}
\IEEEauthorblockA{\textit{Imperial College London}\\
London, UK \\
g.bisbas18@imperial.ac.uk}
\and
\IEEEauthorblockN{6\textsuperscript{th} Jan Thorbecke}
\IEEEauthorblockA{\textit{TU-Delft}\\
Delft, NL \\
J.W.Thorbecke@tudelft.nl}
\and
\IEEEauthorblockN{7\textsuperscript{th} Felix J. Herrmann}
\IEEEauthorblockA{\textit{Georgia Institute of Technology}\\
Atlanta, GA\\
felix.herrmann@gatech.edu}
\and
\IEEEauthorblockN{8\textsuperscript{th} Gerard J. Gorman}
\IEEEauthorblockA{\textit{Imperial College London}\\
London, UK \\
g.gorman@imperial.ac.uk}
}

@@

These 21 fields, for the 5.9 billion point grid defined above, lead to a
minimum memory requirement of 461Gb for modelling alone. For this experiment,
access was obtained for small HPC VMs (on Azure) called \texttt{Standard\_H16r}.
These VMs contain 16 core Intel Xeon E5 2667 v3 chips, with no hyperthreading,
and 32 nodes were used for a single source experiment
(i.e. a single wave-equation was solved). We
used a 12th order discretization in space that led to 2.8TFLOP/time-step being
computed by this model and the elastic wave was propagated for 16 seconds
(23000 time steps). Completion of this modelling run took 16 hours, converting to
1.1TFLOP/s. While these numbers may appear low, it should be noted that
the elastic kernel is extremely memory bound, while the TTI kernel is nearly
compute bound (see rooflines in \cite{louboutin2016ppf, devito-api, devito-compiler})
making it more computationally efficient, particularly in combination
with MPI. Future work will involve working on InfiniBand enabled and
true HPC VMs on Azure to achieve Cloud performance on par with that of state of
the art HPC clusters. Extrapolating from the performance obtained in this
experiment, and assuming a fairly standard setup of 5000 independent
source experiments, computing an elastic synthetic dataset would require 322
EFLOPs (23k time-steps x 2.8TFLOP/time-step x 500
sources), or utilizing the full scalability and computing all sources in
parallel this becomes 5.5PFLOP/s.

>>>>

\title{Scaling through abstractions – high-performance vectorial wave simulations for seismic inversion with Devito}

\author{
\IEEEauthorblockN{1\textsuperscript{st} Mathias Louboutin}
\IEEEauthorblockA{\textit{Georgia Institute of Technology}\\
Atlanta, GA \\
mlouboutin3@gatech.edu}
\and
\IEEEauthorblockN{2\textsuperscript{nd} Fabio Luporini}
\IEEEauthorblockA{\textit{Devito Codes}\\
London, UK \\
fabio@devitocodes.com}
\and
\IEEEauthorblockN{3\textsuperscript{rd} Rhodri Nelson}
\IEEEauthorblockA{\textit{Imperial College London}\\
London, UK \\
rhodri.nelson@imperial.ac.uk}
\and
\IEEEauthorblockN{4\textsuperscript{th} Philipp Witte}
\IEEEauthorblockA{\textit{Georgia Institute of Technology}\\
Atlanta, GA \\
pwitte3@gatech.edu}
\and
\IEEEauthorblockN{5\textsuperscript{th} George Bisbas}
\IEEEauthorblockA{\textit{Imperial College London}\\
London, UK \\
g.bisbas18@imperial.ac.uk}
\and
\IEEEauthorblockN{6\textsuperscript{th} Jan Thorbecke}
\IEEEauthorblockA{\textit{TU-Delft}\\
Delft, NL \\
J.W.Thorbecke@tudelft.nl}
\and
\IEEEauthorblockN{7\textsuperscript{th} Felix J. Herrmann}
\IEEEauthorblockA{\textit{Georgia Institute of Technology}\\
Atlanta, GA\\
felix.herrmann@gatech.edu}
\and
\IEEEauthorblockN{8\textsuperscript{th} Gerard J. Gorman}
\IEEEauthorblockA{\textit{Imperial College London}\\
London, UK \\
g.gorman@imperial.ac.uk}
}

% \author{\IEEEauthorblockN{1\textsuperscript{st} Given Name Surname}
% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
% \textit{name of organization (of Aff.)}\\
% City, Country \\
% email address or ORCID}
% \and
% \IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
% \textit{name of organization (of Aff.)}\\
% City, Country \\
% email address or ORCID}
% \and
% \IEEEauthorblockN{3\textsuperscript{rd} Given Name Surname}
% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
% \textit{name of organization (of Aff.)}\\
% City, Country \\
% email address or ORCID}
% \and
% \IEEEauthorblockN{4\textsuperscript{th} Given Name Surname}
% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
% \textit{name of organization (of Aff.)}\\
% City, Country \\
% email address or ORCID}
% \and
% \IEEEauthorblockN{5\textsuperscript{th} Given Name Surname}
% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
% \textit{name of organization (of Aff.)}\\
% City, Country \\
% email address or ORCID}
% \and
% \IEEEauthorblockN{6\textsuperscript{th} Given Name Surname}
% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
% \textit{name of organization (of Aff.)}\\
% City, Country \\
% email address or ORCID}
% }

@@

These 21 fields, for the 5.9 billion point grid defined above, lead to a
minimum memory requirement of 461Gb for modelling alone. For this experiment,
access was obtained for small HPC VMs (on Azure) called \texttt{Standard\_H16r}.
These VMs contain 16 core Intel Xeon E5 2667 v3 chips, with no hyperthreading,
and 32 nodes were used for a single source experiment
(i.e. a single wave-equation was solved). We
used a 12th order discretization in space that led to 2.8TFLOP/time-step being
computed by this model and the elastic wave was propagated for 16 seconds
(23000 time steps). Completion of this modelling run took 16 hours, converting to
1.1TFLOP/s. While these numbers may appear low, it should be noted that
the elastic kernel is extremely memory bound, while the TTI kernel is nearly
compute bound (see rooflines in \cite{louboutin2016ppf, devito-api, devito-compiler})
making it more computationally efficient, particularly in combination
with MPI. Future work will involve working on InfiniBand enabled and
true HPC VMs on Azure to achieve Cloud performance on par with that of state of
the art HPC clusters. Extrapolating from the performance obtained in this
experiment, and assuming a fairly standard setup of 5000 independent
source experiments, computing an elastic synthetic dataset would require 322
EFLOPs (23k time-steps x 2.8TFLOP/time-step x 5000
sources), or utilizing the full scalability and computing all sources in
parallel this becomes 5.5PFLOP/s.
